Containers and R---the Rockerverse and beyond

### 2019-11-06 manuscript.Rmd
add some intro sentences, fix docker example command

## Clients
<!-- first order packages, talking directyl to Docker -->

@@

## Capture and create environments [@nuest]
<!-- second order packages -->

@@

## Deployment to cloud services [@nuest, @richfitz, @rcannood]

- [`babelwhale`](https://cran.r-project.org/web/packages/babelwhale/index.html): Running a Docker from R with Singularity or Docker as back-end. This is really useful in HPC environments where a user might not have root access but is able to install Singularity instead of Docker. [@rcannood]
- `RSelenium`
- [`googleComputeEngineR`](https://cloudyr.github.io/googleComputeEngineR/) (function `docker_run`)
  - I don't think `docker_run` is main focus for containers for this package, its more how it uses Docker to launch custom environments that can be templated (e.g. rstudio, shiny etc.) and for parrallisation e.g. `library(future)`  using GCE VMs as endpoints. [@MarkEdmondson1234]
- [`analogsea`](https://github.com/sckott/analogsea) (digital ocean R client)

@@

- R-Hub
- DevOps
  - https://www.opencpu.org/posts/opencpu-with-docker/

@@

[RCloud](https://github.com/att/rcloud/tree/master/docker) uses a `rocker/drd` base image for creation of collaborative data analysis and visualisation environments.

## Deployment, processing, cloud [div]

- Docker images for cloud services [@MarkEdmondson1234]
  - the most popular functionality for `googleComputeEngineR` is its use of Docker to enable parallel processing across VMS - [some demos here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html)
- Google Cloud Run - CaaS (Containers as a Service) that lets you launch a Docker container without worrying about underlying infrastructure. An R implementation is shown here at [cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR) which uses it to create a scalable R plumber API. 
- [`plumber`](https://www.rplumber.io/docs/hosting.html#docker)
- `batchtools` \citep{Lang2017batchtools} can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
- scalable deployments, e.g. start with numerous Shiny talks mentioning Rocker at useR!2017
- [dynmethods](https://github.com/dynverse/dynmethods) [@rcannood]: In order to evaluate ±50 computational methods which all used different environments (R, Python, C++, ...), we wrapped each of them in a docker container and can execute these methods from R. Again, all of these containers are being built on travis-ci, and will only be pushed to docker hub if the integration test succeeds.
- [ShinyProxy](https://www.shinyproxy.io/) [creates a container](https://github.com/openanalytics/shinyproxy/blob/master/src/main/java/eu/openanalytics/services/DockerService.java#L388) for each user
- [production deployment of neural networks](https://github.com/tmobile/r-tensorflow-api) [@jnolis, @nolistic]

@@

# Other containerisation platforms [NN]

- R images for Singularity
- Running Rocker images with
  - Singularity
  - CoreOS rkt?
- nix?

>>>>

## Clients

A number first order R packages exist that talk directly to Docker.
They can be used to control the Docker daemon, e.g. to build images or run containers, from R without explicit knowledge of the Docker command line interface.

@@

## Capture and create environments [@nuest]

Several second order R packages attempt to make the process of creating Docker images and using containers for specific tasks, such as running tests or rendering reproducible reports, easier.

@@

## Deployment to cloud services [div]

The cloud is the natural environment of containers, and becomes the go-to mechanism to expose applications written in R to users.

- [`babelwhale`](https://cran.r-project.org/web/packages/babelwhale/index.html): Running a Docker from R with Singularity or Docker as back-end. This is really useful in HPC environments where a user might not have root access but is able to install Singularity instead of Docker. [@rcannood]
- `RSelenium`
- [`googleComputeEngineR`](https://cloudyr.github.io/googleComputeEngineR/) (function `docker_run`)
  - I don't think `docker_run` is main focus for containers for this package, its more how it uses Docker to launch custom environments that can be templated (e.g. rstudio, shiny etc.) and for parrallisation e.g. `library(future)`  using GCE VMs as endpoints. [@MarkEdmondson1234]
- [`analogsea`](https://github.com/sckott/analogsea) (digital ocean R client)
- [`plumber`](https://www.rplumber.io/docs/hosting.html#docker)
- [production deployment of neural networks](https://github.com/tmobile/r-tensorflow-api) [@jnolis, @nolistic]

@@

The controlled nature of containers, i.e. it is possible to define the software environment very well, even on remote machines, make them also useful for continuous integration (CI) and continuous delivery of applications.

- DevOps
  - https://www.opencpu.org/posts/opencpu-with-docker/

@@

- [RCloud](https://github.com/att/rcloud/tree/master/docker) uses a `rocker/drd` base image for creation of collaborative data analysis and visualisation environments.
- [ShinyProxy](https://www.shinyproxy.io/) [creates a container](https://github.com/openanalytics/shinyproxy/blob/master/src/main/java/eu/openanalytics/services/DockerService.java#L388) for each user

## Processing [div]

The portability of containers becomes particularly useful when complex processing tasks shall be offloaded to a server.

- Docker images for cloud services [@MarkEdmondson1234]
  - the most popular functionality for `googleComputeEngineR` is its use of Docker to enable parallel processing across VMS - [some demos here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html)
- Google Cloud Run - CaaS (Containers as a Service) that lets you launch a Docker container without worrying about underlying infrastructure. An R implementation is shown here at [cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR) which uses it to create a scalable R plumber API. 
- `batchtools` \citep{Lang2017batchtools} can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
- scalable deployments, e.g. start with numerous Shiny talks mentioning Rocker at useR!2017
- [dynmethods](https://github.com/dynverse/dynmethods) [@rcannood]: In order to evaluate ±50 computational methods which all used different environments (R, Python, C++, ...), we wrapped each of them in a docker container and can execute these methods from R. Again, all of these containers are being built on travis-ci, and will only be pushed to docker hub if the integration test succeeds.

@@

- **[R-Hub](https://github.com/r-hub/)**

# Other containerisation platforms

Docker is not the only containerisation software.
An alternative stemming from the domain of high-perfomance computing is Singularity \citep{kurtzer_singularity_2017}.
Singularity can run Docker images, and in the case of Rocker works out of the box if the main process is R, e.g., in `rocker/r-base`, but does not succeed in running images where there is an init script, e.g. in containers that by default run RStudio.
In the latter case, a `Singularity` file, a set of instructions akin to a `Dockerfile` needs to be used.
